An essentially algebraic formulation of dependent type theory

### 2015-06-16 weakeningcats.tex
E'-systems

\section{The definition of E-systems}\label{sec:esys_defn}
\subsection{Weakening systems}

@@

\subsection{E-systems}

@@

\subsection{Internal E-systems}

\subsubsection{Internal weakening systems}

\subsubsection{Internal projection systems}

\subsubsection{Internal substitution systems}

\subsubsection{Internal E-systems}

@@

\subsection{From D-systems to E-systems}

@@

\begin{defn}
A \emph{setup for categories with families} consists of
\begin{enumerate}
\item A category $\cat{C}$ with a terminal object $1$.
\item A presheaf $\mathrm{Ty}:\op{\cat{C}}\to\mathbf{Set}$. We will write
$A[\gamma]$ for $\mathrm{Ty}(\gamma,A)$. 
\item A commuting triangle
\begin{equation*}
\begin{tikzcd}
\int_\cat{C}\mathrm{Ty} \arrow[dr,swap,"\pi"] \arrow[rr,"p"] & & \cat{C}^{\to} \arrow[dl,"\mathrm{cod}"] \\
& \cat{C}
\end{tikzcd}
\end{equation*} 
We will write $p_A:\ctxext{\Gamma}{A}\to\Gamma$
for the morphism $U(\Gamma,A)$ into $\Gamma$, for each object $\Gamma\in\cat{C}$ and
$A\in\mathrm{Ty}(\Gamma)$.
\end{enumerate}
\end{defn}

\begin{defn}
A \emph{category with families} consists of a setup with
\begin{enumerate}
\item[(iv)] a preasheaf $\mathrm{Tm}:\op{(\int_\cat{C}\mathrm{Ty})}\to\mathbf{Set}$,
\item[(v)] a term $q_A\in\mathrm{Tm}(\ctxext{\Gamma}{A},A[p_A])$ for every
$A\in\mathrm{Ty}(\Gamma)$, 
\end{enumerate}
such that for every $\gamma:\Delta\to\Gamma$ in $\cat{C}$ and 
$a\in \mathrm{Tm}(\Delta,A[\gamma])$ there is a unique morphism 
$\theta:\Delta\to \ctxext{\Gamma}{A}$ such that $p_A\circ\theta\jdeq\gamma$ and 
$q_A[\theta]\jdeq a$.
\end{defn}

\begin{lem}
Let $\mathbb{D}$ be a category with families. Then there are bijections
\begin{equation*}
\mathrm{Tm}(\Delta,A[\gamma])\cong\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq \gamma\}
\end{equation*}
natural in $\gamma$ and $A$.
\end{lem}

\begin{proof}
Let $\theta:\Delta\to\ctxext{\Gamma}{A}$ be such that $p_A\circ\theta\jdeq\gamma$. 
Then $\subst{\theta}{q_A}$ is a term
of $\subst{\theta}{{p_A}{A}}$. By the functoriality of $\mathrm{Ty}$, we have
$\subst{\theta}{{p_A}{A}}\jdeq \subst{p_A\circ\theta}{A}\jdeq\subst{\gamma}{A}$. 
So we obtain $\subst{\theta}{q_A}\in\mathrm{Tm}(\Delta,\subst{\gamma}{A})$. Now
the uniqueness of $\theta$ implies that the map $\theta\mapsto\subst{\theta}{q_A}$
is a bijection. 
\end{proof}

@@

\begin{defn}
A \emph{D-system} consists of a setup for categories with families
such that for every $\gamma:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, there
is a bijection
\begin{equation*}
\{\eta:\Delta\to\ctxext{\Delta}{\subst{\gamma}{A}}\mid p_{\subst{\gamma}{A}}\circ\eta\jdeq\catid{\Delta}\}
  \cong
\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq\gamma\}
\end{equation*}
natural in $\gamma$ and $A$.
\end{defn}

@@

\begin{defn}
There is a functor $U$ from D-systems to locally stratified categories.
\end{defn}

\begin{constr}
Let $\mathbb{D}$ be a D-system with underlying category $\cat{C}$. We define
$U(\mathbb{D})$ to be the category with the same objects as $\cat{C}$, where the
morphisms are generated by the projections $p_A:\ctxext{\Gamma}{A}\to\Gamma$.

We have to prove that this category is locally stratified. Let $\Gamma\in\cat{C}$. 
\end{constr}

\begin{defn}
For any D-system $\mathbb{D}$, $U(\mathbb{D})$ can be given the structure of
an E-system without terminal object. Thus, we get a functor
$\mathbf{Dsys}\to\mathbf{Esys}_{\circ}$
\end{defn}

\begin{constr}
We first need to define a term structure on $U(\mathbb{D})$. 
Next, we define a weakening structure on $\cat{C}_p$. Since the morphisms of
$\cat{C}_p$ are composites of projections, we will define
\begin{equation*}
W_{p_{A_n}\circ\cdots\circ p_{A_1}}\defeq W_{p_{A_n}}\circ\cdots W_{p_{A_1}}.
\end{equation*}
Therefore, we only have to define functors 
$W_A:\cat{C}_p/\Gamma\to\cat{C}_p/\ctxext{\Gamma}{A}$ with term structure for every 
$A\in\mathrm{Ty}(\Gamma)$. Since morphisms are composites of projections, we
can define this functor by induction. Let $W_A(\catid{\Gamma})\defeq
\catid{\ctxext{\Gamma}{A}}$ and $W_A(p_Q\circ f)\defeq$
\end{constr}

@@

\section{An alternative characterization of E-systems}

\begin{defn}
An E'-system consists of a category $\cat{C}$ with a subcategory $\cat{F}$ with the
same objects as $\cat{C}$, with
the following structure
\begin{enumerate}
\item $\cat{F}$ has a terminal object.
\item For any $f:\Delta\to\Gamma$ in $\cat{C}$, a functor $f_\ast:\cat{F}/\Gamma\to
\cat{F}/\Delta$ %and
%for any $Q\in\cat{F}/\ctxext{\Gamma}{B}$ and any section $g$ of $Q$ a section
%$f_\ast(g)$ of $f_\ast(Q)$, 
so that $f_\ast$ is contravariant in $f$ and preserves the terminal object. 
\item For any $A\in\cat{F}/\Gamma$ and $f:\Delta\to\Gamma$, factorizations of
$f$ through $A$ are naturally the same thing as sections of
$f_\ast(A)$, i.e. there is an isomorphism
\begin{equation*}
\zeta_{A,f}%\defeq\theta\mapsto f_\ast(\theta)
:\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\}
  \cong
\{\eta:\Delta\to\ctxext{\Delta}{(f_\ast(A))}\mid f_\ast(A)\circ\eta\jdeq\catid{\Delta}\}
\end{equation*}
in the presheaf category 
$\mathrm{Psh}(\op{(\cat{F}/\Gamma)}\times(\cat{C}/\Gamma))$. 
\end{enumerate}
\end{defn}

\begin{rmk}
Explicitly, the
naturality condition states that for any $P\in\cat{F}/\ctxext{\Gamma}{A}$,
$g:\greek{E}\to\Delta$ and $\theta:\Delta\to\ctxext{{\Gamma}{A}}{P}$ such that
$P\circ\theta\jdeq f$, we have the equality
\begin{equation*}
\zeta_{A,f\circ g}(P\circ\theta\circ g)\jdeq f^\ast(P)\circ\zeta_{P,f}(\theta)\circ g.
\end{equation*}
\end{rmk}

@@

\begin{lem}
Let $\mathbb{E}$ be an E'-system. Then there is a function $T(A)\to T(f_\ast(A))$
for any $f:\Delta\to\Gamma$ in $\cat{C}$ and $A\in\cat{F}/\Gamma$, contravariant
in $f$. 
\end{lem}

@@

\begin{defn}
Consider $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $f:\Delta\to\Gamma$,
$\theta_1:\Delta\to\ctxext{\Gamma}{A}$ such that $A\circ\theta_1\jdeq f$, 
and $\theta_2:\Delta\to\ctxext{{\Gamma}{A}}{P}$ such that $P\circ\theta_2\jdeq\theta_1$.  
\end{defn}

\begin{defn}
For any $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $f:\Delta\to\Gamma$ 
there is a morphism $\tilde{f}:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}$
with the property that $A\circ\tilde{f}\jdeq f\circ f_\ast(A)$, and
we have isomorphisms
\begin{align*}
\{\theta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \tilde{f}\}
&  \cong
\{\eta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{(f_\ast(A))}}{f_\ast(P)}\mid f_\ast(P)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\}
\end{align*}
\end{defn}

\begin{constr}
\begin{align*}
& \{ \theta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\circ f_\ast(A) \} \\
  & \cong
\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(f\circ f_\ast(A))_\ast(A)}
  \mid (f\circ f_\ast(A))_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
  & \jdeq
\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(f_\ast(A))_\ast(f_\ast(A))}
  \mid (f\circ f_\ast(A))_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
  & \jdeq
T(W_{f_\ast(A)}(f_\ast(A))).
\end{align*}
Since we have $\tfid{f_\ast(A)}\in T(W_{f_\ast(A)}(f_\ast(A)))$, we obtain 
\begin{equation*}
\tilde{f}\defeq \zeta_{A,f\circ f_\ast(A)}^{-1}(\tfid{f_\ast(A)}):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}.
\end{equation*}
It remains to prove the asserted isomorphism. 
\begin{align*}
\{\theta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Gamma}{A}}{P}\mid P\circ\theta\jdeq \tilde{f}\}
&  \cong
\{\eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{\tilde{f}_\ast(P)} \mid
  \tilde{f}_\ast(P)\circ\eta\jdeq\catid{}\}
  \\
  & \cong
\{\eta:\ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{(f_\ast(A))}}{f_\ast(P)}\mid f_\ast(P)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\}
\end{align*}
\end{constr}

>>>>

\section{E-systems}\label{sec:esys_defn}

\subsection{Weakening systems}

@@

\subsection{The definition of E-systems}

@@

@@

\subsection{From D-systems to locally stratified E'-systems}

@@

\begin{defn}
A \emph{setup for categories with families} consists of
\begin{enumerate}
\item A category $\cat{C}$ with a terminal object $1$.
\item A presheaf $\mathrm{Ty}:\op{\cat{C}}\to\mathbf{Set}$. We will write
$f_\ast(A)$ for $\mathrm{Ty}(f,A)$. 
\item A commuting triangle
\begin{equation*}
\begin{tikzcd}
\int_\cat{C}\mathrm{Ty} \arrow[dr,swap,"\pi"] \arrow[rr,"p"] & & \cat{C}^{\to} \arrow[dl,"\mathrm{cod}"] \\
& \cat{C}
\end{tikzcd}
\end{equation*} 
We will write $p_A:\ctxext{\Gamma}{A}\to\Gamma$
for the morphism $U(\Gamma,A)$ into $\Gamma$, for each object $\Gamma\in\cat{C}$ and
$A\in\mathrm{Ty}(\Gamma)$.
\end{enumerate}
\end{defn}

\begin{defn}
A \emph{category with families} consists of a setup with
\begin{enumerate}
\item[(iv)] a preasheaf $\mathrm{Tm}:\op{(\int_\cat{C}\mathrm{Ty})}\to\mathbf{Set}$.
The action of a morphism $f:(\Delta,f_\ast(A))\to(\Gamma,A)$ of $\int_\cat{C}\mathrm{Ty}$
on a term $x\in \mathrm{Tm}(A)$ is also written by $f_\ast(x)$.
\item[(v)] a term $q_A\in\mathrm{Tm}(\ctxext{\Gamma}{A},(p_A)_\ast(A))$ for every
$A\in\mathrm{Ty}(\Gamma)$, 
\end{enumerate}
such that for every $f:\Delta\to\Gamma$ in $\cat{C}$ and 
$x\in \mathrm{Tm}(\Delta,f_\ast(A))$ there is a unique morphism 
$\tilde{x}:\Delta\to \ctxext{\Gamma}{A}$ such that $p_A\circ\tilde{x}\jdeq f$ and 
$\tilde{x}_\ast(q_A)\jdeq x$.
\end{defn}

\begin{lem}\label{lem:cwf_to_dsys}
Let $\mathbb{D}$ be a category with families. Then there are bijections
\begin{equation*}
\zeta_{A,f}:\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}\cong \mathrm{Tm}(\Delta,f_\ast(A))
\end{equation*}
coherent in $f$ in the sense that
\begin{align*}
\zeta_{A,\catid{\Gamma}}(\theta) & \jdeq \theta \\
\zeta_{A,f\circ g}(\theta) & \jdeq \zeta_{f_\ast(A),g}(\zeta_{A,f}(\theta)\circ g)
\end{align*}
for any $g:X\to\Delta$. 
\end{lem}

\begin{proof}
Let $\theta:\Delta\to\ctxext{\Gamma}{A}$ be such that $p_A\circ\theta\jdeq f$. 
Then $\theta_\ast(q_A)$ is an element
of $\mathrm{Tm}(\theta_\ast((p_A)_\ast(A)))$. By the functoriality of $\mathrm{Ty}$, we have
$\theta_\ast((p_A)_\ast(A))\jdeq (p_A\circ\theta)_\ast(A)\jdeq f_\ast(A)$. 
So we obtain $\theta_\ast(q_A)\in\mathrm{Tm}(\Delta,f_\ast(A))$. Now
the uniqueness of $\theta$ implies that the map $\theta\mapsto \theta_\ast(q_A)$
is a bijection.

It remains to verify the coherence identities. Let $\theta : \Gamma\to\ctxext{\Gamma}{A}$
be such that $p_A\circ\theta\jdeq\catid{\Gamma}$. We wish to show that
$\theta_\ast(q_A)\jdeq\theta$.  
\begin{align*}
\theta_\ast(q_A)
& \jdeq \theta_\ast(\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})) \\
& \jdeq \cdots \\
& \jdeq \theta
\end{align*}
\end{proof}

@@

\begin{defn}
A \emph{D-system} consists of a setup for categories with families
such that for every $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, there
is a bijection
\begin{equation*}
\zeta_{A,f}:\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid p_A\circ\theta\jdeq f\}
  \cong
\{\eta:\Delta\to\ctxext{\Delta}{f_\ast(A)}\mid p_{f_\ast(A)}\circ\eta\jdeq\catid{\Delta}\}
\end{equation*}
natural in $f$ in the sense that the equalities
\begin{align*}
\zeta_{A,\catid{\Gamma}}(\theta) & \jdeq \theta \\
\zeta_{A,f\circ g}(\theta\circ g) & \jdeq \zeta_{f_\ast(A),g}(\zeta_{A,f}(\theta)\circ g)
\end{align*}
hold for any $f:\Delta\to\Gamma$ and $g:X\to\Delta$, and any $\theta:\Delta\to\ctxext{\Gamma}{A}$
such that $p_A\circ\theta\jdeq f$. 
\end{defn}

@@

\begin{proof}
In \autoref{lem:cwf_to_dsys} we have shown that every category with families is
a D-system. Thus, it suffices to show the converse.

Let $\mathbb{D}$ be a D-system. For any object $\Gamma$ of $\mathbb{D}$ and any
type $A\in\mathrm{Ty}(\Gamma)$, we define 
\begin{equation*}
\mathrm{Tm}(A)\defeq \{ x :\Gamma\to\ctxext{\Gamma}{A}\mid p_A\circ x\jdeq\catid{\Gamma}\}.
\end{equation*}
For any morphism $f:\Delta\to\Gamma$ and $A\in\mathrm{Ty}(\Gamma)$, we define
the function $\mathrm{Tm}(f):\mathrm{Tm}(A)\to\mathrm{Tm}(f_\ast(A))$ by
\begin{equation*}
\mathrm{Tm}(f,x)\defeq \zeta_{A,f}(x\circ f).
\end{equation*}
We have to verify that this satisfies the functorial identities. For $f:\Delta\to\Gamma$
and $g:X\to\Delta$ and any $x\in\mathrm{Tm}(A)$, we have
\begin{align*}
\mathrm{Tm}(f\circ g,x) & \jdeq \zeta_{A,f\circ g}(x\circ f\circ g) \\
& \jdeq \zeta_{f_\ast(A),g}(\zeta_{A,f}(x\circ f)\circ g) \\
& \jdeq \mathrm{Tm}(g,\mathrm{Tm}(f,x)).
\end{align*}
and
\begin{equation*}
\mathrm{Tm}(\catid{\Gamma},x) \jdeq \zeta_{A,\catid{\Gamma}}(x) \jdeq x.
\end{equation*}
This finishes the construction of the presheaf $\mathrm{Tm}$ on $\int_{\cat{C}}\mathrm{Ty}$.

Next, we have to find $q_A\in\mathrm{Tm}((p_A)_\ast(A))$. Note that we have the
isomorphism
\begin{align*}
\mathrm{Tm}((p_A)_\ast(A))
& \jdeq \{\eta :\ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{(p_A)_\ast(A)}\mid p_{(p_A)_\ast(A)}\circ\theta\jdeq\catid{\ctxext{\Gamma}{A}}\} \\
& \cong \{\theta : \ctxext{\Gamma}{A}\to\ctxext{\Gamma}{A} \mid p_A\circ \theta\jdeq p_A\}
\end{align*}
The latter set contains the identity morphism $\catid{\ctxext{\Gamma}{A}}$. Therefore,
we find
\begin{equation*}
q_A \defeq \zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\in\mathrm{Tm}((p_A)_\ast(A)).
\end{equation*}
To verify that we have constructed a category with families, let $f:\Delta\to\Gamma$
and $x\in T(f_\ast(A))$. Note that we have the morphism 
$\tilde{x}\defeq \zeta_{A,f}^{-1}(x):\Delta\to\ctxext{\Gamma}{A}$. 
We need to check that $\tilde{x}$ is unique with the property that 
$\mathrm{Tm}(\tilde{x},q_A)\jdeq x$. The coherence identities can be used to see
that this is indeed the case:
\begin{align*}
\mathrm{Tm}(\tilde{x},q_A)
& \jdeq \zeta_{(p_A)_\ast(A),\tilde{x}}(q_A\circ\tilde{x}) \\
& \jdeq \zeta_{(p_A)_\ast(A),\tilde{x}}(\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\circ\tilde{x}) \\
& \jdeq \zeta_{A,p_A\circ\tilde{x}}(\tilde{x}) \\
& \jdeq \zeta_{A,f}(\tilde{x}) \\
& \jdeq x.
\end{align*}
For the uniqueness, let $\theta:\Delta\to\ctxext{\Gamma}{A}$ be any morphism such that
$p_A\circ\theta\jdeq f$ and $\mathrm{Tm}(\theta,q_A)\jdeq x$. Then we have
\begin{align*}
x & \jdeq \mathrm{Tm}(\theta,q_A) \\
& \jdeq \zeta_{(p_A)_\ast(A),\theta}(q_A\circ\theta) \\
& \jdeq \zeta_{(p_A)_\ast(A),\theta}(\zeta_{A,p_A}(\catid{\ctxext{\Gamma}{A}})\circ\theta \\
& \jdeq \zeta_{A,p_A\circ\theta}(\theta) \\
& \jdeq \zeta_{A,f}(\theta).
\end{align*}
Therefore we see that $\tilde{x}\jdeq\zeta_{A,f}^{-1}(x)\jdeq\theta$.
\end{proof}

\begin{defn}
An E'-system is said to be locally stratified, if the category $\cat{F}$ is
locally stratified.
\end{defn}

\begin{comment}
\begin{rmk}
In the following definition, one could take the underlying category of the
E-system $\mathcal{E}(\mathbb{D})$ associated to a D-system $\mathbb{D}$ to
be the subcategory generated by the projections $p_A$, and construct the
rest of the structure of E-systems on top of that. However, it is not so easy
to characterize the image of the functor $\mathcal{E}$ in that case. Intuitively,
D-systems should correspond to locally stratified E-systems. To be able to prove
such a theorem, we've gone with a slightly modified definition of the underlying
category.
\end{rmk}
\end{comment}

\begin{defn}
For any D-system $\mathbb{D}$ there is 
a locally stratified E'-system $U(\mathbb{D})$ without terminal object in 
$\cat{F}$. Thus, we get a functor $\mathbf{Dsys}\to\mathbf{E'sys}_{\circ}$
\end{defn}

\begin{constr}
Let $\mathbb{D}$ be a D-system with underlying category $\cat{C}$. We define
$\cat{F}\subseteq\cat{C}$ to be the
\begin{comment}
category consisting of 
\begin{enumerate}
\item equivalence classes of finite lists of the form $(\Gamma,A_1,\ldots A_n)$, where $\Gamma$ is
an object of $\cat{C}$ and $A_{i+1}\in\mathrm{Ty}(\ctxext{{{\Gamma}{A_1}}{\cdots}}{A_i})$, where
the equivalence relation is generated by
\begin{equation*}
(\ctxext{{{\Gamma}{A_1}}{\cdots}}{A_n},A_{n+1},\ldots,A_{n+m})\sim (\Gamma,A_1,\ldots,A_{n+m}).
\end{equation*}
\item
\end{enumerate}
\end{comment}
subcategory with the same objects as
$\cat{C}$, where the morphisms are generated by the projections $p_A$.

Let $f:\Delta\to\Gamma$. We define $f_\ast:\cat{F}/\Gamma\to\cat{F}/\Delta$ by
induction on the length of the arrow $A$ into $\Gamma$, simultaneously with a
morphism $\pi_2(f,A)$ from $\ctxext{\Delta}{f_\ast(A)}$ into $\ctxext{\Gamma}{A}$
for which the square
\begin{equation*}
\begin{tikzcd}
\ctxext{\Delta}{f_\ast(A)} \arrow[r,"{\pi_2(f,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
\Delta \arrow[r,swap,"f"] & \Gamma
\end{tikzcd}
\end{equation*}
is a pullback square. We define
$f_\ast(\catid{\Gamma})\defeq\catid{\Delta}$, and $\pi_2(f,\catid{\Gamma})\defeq f:\ctxext{\Delta}{f_\ast(\catid{\Gamma})}\to\ctxext{\Gamma}{\catid{\Gamma}}$. The resulting square is trivially
a pullback square. Now suppose we have a pullback square as above.
and let $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then we define
\begin{equation*}
f_\ast(A\circ p_B)\defeq \pi_2(f,A)_\ast(B).
\end{equation*}
By \autoref{lem:EE_pb}, we get the top pullback square in the diagram
\begin{equation*}
\begin{tikzcd}[column sep=6em]
\ctxext{{\Delta}{f_\ast(A)}}{f_\ast(B)} \arrow[d,fib] \arrow[r,"{\pi_2(\pi_2(f,A),B)}"] & \ctxext{{\Gamma}{A}}{B} \arrow[d,fib] \\
\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
\Delta \arrow[r,"f"] & \Gamma
\end{tikzcd}
\end{equation*}
Since the bottom square is assumed to be a pullback, we get that the outer rectangle
is a pullback, as desired.

Next, we define a weakening structure on $\cat{C}_p$. Since the morphisms of
$\cat{C}_p$ are composites of projections, we will define
\begin{equation*}
W_{p_{A_n}\circ\cdots\circ p_{A_1}}\defeq W_{p_{A_n}}\circ\cdots W_{p_{A_1}}.
\end{equation*}
Therefore, we only have to define functors 
$W_A:\cat{C}_p/\Gamma\to\cat{C}_p/\ctxext{\Gamma}{A}$ with term structure for every 
$A\in\mathrm{Ty}(\Gamma)$. Since morphisms are composites of projections, we
can define this functor by induction. Let $W_A(\catid{\Gamma})\defeq
\catid{\ctxext{\Gamma}{A}}$ and $W_A(p_Q\circ f)\defeq$
\end{constr}

@@

\subsection{The equivalence of E-systems and E'-systems}

The definition of D-system suggests a more general structure related to E-systems.

\begin{defn}
An \emph{E'-system} consists of a category $\cat{C}$ with a subcategory $\cat{F}$ with the
same objects as $\cat{C}$, with
the following structure
\begin{enumerate}
\item $\cat{F}$ has a terminal object.
\item For any $f:\Delta\to\Gamma$ in $\cat{C}$, a functor $f_\ast:\cat{F}/\Gamma\to
\cat{F}/\Delta$ %and
%for any $Q\in\cat{F}/\ctxext{\Gamma}{B}$ and any section $g$ of $Q$ a section
%$f_\ast(g)$ of $f_\ast(Q)$, 
so that $f_\ast$ is contravariant in $f$ and preserves the terminal object. 
\item For any $A\in\cat{F}/\Gamma$ and $f:\Delta\to\Gamma$, lifts of
$f$ to $A$ are naturally the same thing as sections of
$f_\ast(A)$, i.e. there is an isomorphism
\begin{equation*}
\zeta_{A,f}%\defeq\theta\mapsto f_\ast(\theta)
:\{\theta:\Delta\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\}
  \cong
\{\eta:\Delta\to\ctxext{\Delta}{(f_\ast(A))}\mid f_\ast(A)\circ\eta\jdeq\catid{\Delta}\}
\end{equation*}
which are natural in $A$ and $f$, in the sense that
\begin{enumerate}
\item For any $P\in\cat{F}/\ctxext{\Gamma}{A}$, $f:\Delta\to\Gamma$ and
$\theta:\Delta\to\ctxext{{\Gamma}{A}}{P}$ such that $(\ctxext{A}{P})\circ\theta\jdeq f$,
one hase
\begin{equation*}
\zeta_{A,f}(P\circ\theta)\jdeq f_\ast(P)\circ \zeta_{\ctxext{A}{P},f}(\theta).
\end{equation*}
\item For any $A\in\cat{F}/\Gamma$, $f:\Delta\to\Gamma$, $g:X\to\Delta$
and $\theta:\Delta\to\ctxext{\Gamma}{A}$ such that $A\circ\theta\jdeq f$, one
has
\begin{align*}
\zeta_{A,\catid{\Gamma}}(\theta) & \jdeq \theta \\
\zeta_{A,f\circ g}(\theta\circ g) & \jdeq\zeta_{f_\ast(A),g}(\zeta_{A,f}(\theta)\circ g).
\end{align*}
\end{enumerate}
\end{enumerate}
\end{defn}

@@

\begin{lem}\label{lem:ptm}
Let $\mathbb{E}$ be an E'-system. Then there is a function $T(A)\to T(f_\ast(A))$
for any $f:\Delta\to\Gamma$ in $\cat{C}$ and $A\in\cat{F}/\Gamma$, contravariant
in $f$. 
\end{lem}

@@

\begin{defn}
For any $A\in\cat{F}/\Gamma$, we define
$W_A\defeq A_\ast:\cat{F}/\Gamma\to\cat{F}/\ctxext{\Gamma}{A}$. Then we get
for any $A,B\in\cat{F}/\Gamma$ an isomorphism
\begin{equation*}
\mathrm{hom}_{\cat{C}/\Gamma}(A,B)
  \cong
T(W_A(B)).
\end{equation*}
In particular, we get a term $\idtm{A}\in T(W_A(A))$. 
\end{defn}

\begin{constr}
Let $A,B\in\cat{F}/\Gamma$. Then we have the isomorphism
\begin{align*}
T(W_A(B)) & \jdeq \{\eta:\ctxext{\Gamma}{A}\to\ctxext{{\Gamma}{A}}{W_A(B)}\mid W_A(B)\circ\eta\jdeq\catid{\ctxext{\Gamma}{A}}\} \\
  & \cong
\{\theta:\ctxext{\Gamma}{A}\to\ctxext{\Gamma}{B}\mid B\circ\theta\jdeq A\},
\end{align*}
so we find $\idtm{A}\defeq\zeta_{A,A}(\catid{\ctxext{\Gamma}{A}})\in T(W_A(A))$.
\end{constr}

\begin{defn}
For any $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $f:\Delta\to\Gamma$ 
there is a morphism $\pi_2(f,A):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}$
such that $A\circ \pi_2(f,A)\jdeq f\circ f_\ast(A)$.
\end{defn}

\begin{constr}
\begin{align*}
& \{ \theta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}\mid A\circ\theta\jdeq f\circ f_\ast(A) \} \\
  & \cong
\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(f\circ f_\ast(A))_\ast(A)}
  \mid (f\circ f_\ast(A))_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
  & \jdeq
\{ \eta : \ctxext{\Delta}{f_\ast(A)}\to\ctxext{{\Delta}{f_\ast(A)}}{(f_\ast(A))_\ast(f_\ast(A))}
  \mid (f\circ f_\ast(A))_\ast(A)\circ\eta\jdeq\catid{\ctxext{\Delta}{f_\ast(A)}}\} \\
  & \jdeq
T(W_{f_\ast(A)}(f_\ast(A))).
\end{align*}
Since we have $\tfid{f_\ast(A)}\in T(W_{f_\ast(A)}(f_\ast(A)))$, we obtain 
\begin{equation*}
\pi_2(f,A)\defeq \zeta_{A,f\circ f_\ast(A)}^{-1}(\tfid{f_\ast(A)}):\ctxext{\Delta}{f_\ast(A)}\to\ctxext{\Gamma}{A}
\end{equation*}
satisfying $A\circ\pi_2(f,A)\jdeq f\circ f_\ast(A)$.
\end{constr}

\begin{lem}
For any $f:\Delta\to\Gamma$, $g:X\to\Delta$ and $A\in\cat{F}/\Gamma$, one has
\begin{equation*}
\pi_2(f,A)\circ\pi_2(g,f_\ast(A))\jdeq\pi_2(f\circ g,A).
\end{equation*}
\end{lem}

\begin{proof}
It suffices to show that
\begin{equation*}
\zeta_{A,f\circ g\circ g_\ast(f_\ast(A))}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A)))\jdeq\tfid{g_\ast(f_\ast(A))}.
\end{equation*}
Because $g\circ g_\ast(f_\ast(A))\jdeq f_\ast(A)\circ\pi_2(g,f_\ast(A))$, we get
to use the naturality of $\zeta$:
\begin{align*}
& \zeta_{A,f\circ g\circ g_\ast(f_\ast(A))}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A))) \\
& \jdeq
\zeta_{A,f\circ f_\ast(A)\circ\pi_2(g,f_\ast(A))}(\pi_2(f,A)\circ\pi_2(g,f_\ast(A))) \\
& \jdeq
\zeta_{f_\ast(A)_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\zeta_{A,f\circ f_\ast(A)}(\pi_2(f,a))\circ\pi_2(g,f_\ast(A))) \\
& \jdeq
\zeta_{f_\ast(A)_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\tfid{f_\ast(A)}\circ\pi_2(g,f_\ast(A))) \\
& \jdeq
\zeta_{f_\ast(A)_\ast(f_\ast(A)),\pi_2(g,f_\ast(A))}(\zeta_{f_\ast(A),f_\ast(A)}(\catid{\ctxext{\Delta}{f_\ast(A)}})\circ\pi_2(g,f_\ast(A))) \\
& \jdeq
\zeta_{f_\ast(A),f_\ast(A)\circ\pi_2(g,f_\ast(A))}(\catid{\ctxext{\Delta}{f_\ast(A)}}\circ\pi_2(g,f_\ast(A))) \\
& \jdeq
\zeta_{f_\ast(A),f_\ast(A)\circ\pi_2(g,f_\ast(A))}(\pi_2(g,f_\ast(A))) \\
& \jdeq
\zeta_{f_\ast(A),g\circ g_\ast(f_\ast(A))}(\pi_2(g,f_\ast(A)))\\
& \jdeq
\tfid{g_\ast(f_\ast(A))}.\qedhere
\end{align*}
\end{proof}

\begin{lem}
Let $f:\Delta\to\Gamma$ and $A\in\cat{F}/\Gamma$. Then one has 
\begin{equation*}
\pi_2(f,A)\circ\zeta_{A,f}(\theta)\jdeq\theta
\end{equation*}
for any $\theta:\Delta\to\ctxext{\Gamma}{A}$ satisfying $A\circ\theta\jdeq f$.
\end{lem}
